---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r load_data}
activity <- read.csv("activity.csv", stringsAsFactors = FALSE)
```

## What is mean total number of steps taken per day?
```{r steps_per_day}
library(dplyr)
library(ggplot2)
steps_per_day <- summarise(group_by(activity, date), total = sum(steps))
qplot(steps_per_day$total, geom = "histogram")
```
  
Mean: `r mean(steps_per_day$total, na.rm = TRUE)`  
Median: `r median(steps_per_day$total, na.rm = TRUE)`   

## What is the average daily activity pattern?
```{r avg_daily_pattern}
steps_per_interval <- summarise(group_by(activity, interval), avg_steps = mean(steps, na.rm = TRUE))
ggplot(steps_per_interval, aes(interval, avg_steps)) + geom_line() 
```
  
Interval `r steps_per_interval[which.max(steps_per_interval$avg_steps), ]$interval` has the max avg steps of `r steps_per_interval[which.max(steps_per_interval$avg_steps), ]$avg_steps`.  
  
## Imputing missing values
Number of rows with NAs: `r nrow(filter(activity, is.na(steps)))`  
I'm going to immpute missing values by using avg on the interval, join it to a new activity dataset and choose to use the avg of the interval in the case of an NA value.
```{r impute}
avg_interval <- summarise(group_by(activity, interval), avg = mean(steps, na.rm = TRUE))
activity_with_avgs <- left_join(activity, avg_interval, by = c("interval"))
imputed_activity <- mutate(activity_with_avgs, steps = ifelse(is.na(steps), avg, steps))
imputed_activity <- select(imputed_activity, -avg)
imputed_steps_per_day <- summarise(group_by(imputed_activity, date), total = sum(steps))
qplot(imputed_steps_per_day$total, geom = "histogram")

```
  
Mean: `r mean(imputed_steps_per_day$total, na.rm = TRUE)`  
Median: `r median(imputed_steps_per_day$total, na.rm = TRUE)`   
The median settles closer to the mean. The histograms look the same but the bin counts are higher.  

## Are there differences in activity patterns between weekdays and weekends?
```{r weekdays}
imputed_activity_weekdays <- mutate(imputed_activity, weekday = ifelse((weekdays(as.Date(date)) == "Saturday" | weekdays(as.Date(date)) == "Sunday"), "weekend", "weekday"))
imputed_activity_weekdays$weekday <- as.factor(imputed_activity_weekdays$weekday)
steps_per_interval_weekday <- summarise(group_by(imputed_activity_weekdays, interval, weekday), avg_steps = mean(steps, na.rm = TRUE))
ggplot(steps_per_interval_weekday, aes(interval, avg_steps)) + geom_line() + facet_grid(weekday ~ .)
```